### 2020-06-10 report/sections/evalconcl/concl/contributions.tex
Report: progress on conclusion
Report: complete conclusion

\paragraph{Note: from PLACES}

We have presented an MPST-based framework for developing full-stack interactive
TypeScript applications with WebSocket communications.
The implementation conforms to a specified
protocol, statically providing linear channel usage guarantees and affine
channel usage guarantees for back-end and front-end targets respectively.

>>>>

In this project, we develop and present an
MPST-based framework for developing full-stack
interactive TypeScript applications over WebSocket transport.

We motivate our API generation approach from 
\cite{Hybrid2016,PureScript2019} to generate TypeScript APIs
from a Scribble protocol specification.
Developers write their full-stack applications
by implementing the generated APIs, and can enjoy 
communication safety guarantees by construction.
We explain how our session type encodings target the Node.js
server-side runtime and React.js framework in
\cref{chap:node,chap:react} for server-side
and browser-side endpoints respectively.

In particular, we present a novel approach for integrating
session types into web-based GUI programming based on
translating the theory on model types \cite{MVU2020}
to idiomatic practices on the React.js framework.
With respect to session type theory, implementations using
the generated APIs statically enjoy linear channel usage
guarantees and affine channel usage guarantees for back-end
and front-end targets respectively.

Compared to previous work, we are not only able to
\textbf{statically} provide the same level of communication safety
guarantees from multiparty session type theory, but we do so
using modern web programming practices to increase
the relevance and usability of our work in industry:
our work targets TypeScript, Node.js and React.js,
and we explain in \cref{chap:ext} how we support advanced
web development idioms such as asynchronous implementations.

In addition, existing proposals on API generation for web development
over WebSockets \cite{PureScript2019} only support protocols
implementing server-centric topologies, in order to be 
compatible with WebSocket transport.
We formalise a theory of routed multiparty session types in 
\cref{chap:theory} to prove that it is possible to relax the
server-centric topology assumption over WebSocket transport in a way 
that preserves communication along with the communication safety 
properties inherited from canonical session type theory.

% PLACES
Our work has received positive feedback from academia.
The project has been accepted and published in the
Proceedings of the 12th International Workshop on
Programming Language Approaches to Concurrency- \& 
Communication-cEntric Software (PLACES 2020) \cite{PLACES2020}.
We were also invited to present this work as 
\emph{Generating Interactive WebSocket Applications in TypeScript}
at PLACES 2020.

% OVERALL
Overall, we offer an end-to-end solution for integrating
multiparty session type theory into modern full-stack web programming
practices, in a way that actively encourages developers to
design their application with communication safety in mind.
Our formalism of routed multiparty session types enables our
API generation tool to support a wider range of protocols, and
further reveals the potential
for implementing dynamic communication structures over centralised
network topologies.

### 2020-06-10 report/sections/evalconcl/concl/future.tex
Report: complete conclusion

\subsection{Delegation}
Delegation allows channels to be sent between endpoints.
This directly addresses the limitations presented in
\cref{section:evalatm}: in the \tprotocol{ATM} protocol,
when the \trole{Client}
connects to the \trole{Svr},
the \trole{Svr} should be able to send a private channel
for \trole{Client} to perform bank operations,
whilst concurrently accepting new connections.

@@

\subsection{Explicit Connection Actions}
Hu and Yoshida \cite{ExplicitConnections} extends 
multiparty session type theory
with \textit{explicit connection actions} to support protocols 
with optional participants.
Supporting optional participants is a relevant concept
for web applications, particularly for browser-side endpoints.

@@

\subsection{Transport Abstraction}
The MPST-based web development framework presented by 
King et al. \cite{PureScript2019} defines the session runtime
in a way that parameterises the transportation mechanism.
The developer can run the session over different transportation,
provided that the custom transport implements the behaviour
of the \texttt{Transport} type class.

@@

\subsection{Generalised EFSM Encoding as Typestates}
The concept of \textit{typestates} define the interface of
an object to be dependent on its private state, meaning
it can change at runtime.
This is compatible with the EFSM abstraction from the
MPST framework: the permitted methods (i.e. channel actions)
of the EFSM depends on its current state.
In fact, Kouzapas et al. \cite{StMungo} presents a way to
generate typestate specifications for each endpoint of a
Scribble protocol; Gay et al. \cite{ModularST} proposes
an encoding of session types in object-oriented languages
and discuss their approach with respect to typestates.

Future work can explore implementing typestates
in TypeScript. A linear type system would be useful towards 
implementing typestates, but this is lacking in TypeScript.
One interesting possibility for approximating this would
be to combine \textit{decorators} with \textit{transformers}:
the former lets the developer define typestate-related metadata,
whilst the latter can parse the metadata and
transform the source code into another TypeScript program that acts
as an ``intermediate representation'',
such that only implementations that respect the typestate specification
will type-check with its intermediate representation.

>>>>

\paragraph{Delegation}
Delegation allows channels to be sent between endpoints.
This directly addresses the limitations presented in
\cref{section:evalatm}: in the \tprotocol{ATM} protocol,
when the \trole{Client}
connects to the \trole{Svr},
the \trole{Svr} should be able to send a private channel
for \trole{Client} to perform bank operations,
whilst concurrently accepting new connections.

@@

\paragraph{Explicit Connection Actions}
Hu and Yoshida \cite{ExplicitConnections} extends 
multiparty session type theory
with \textit{explicit connection actions} to support protocols 
with optional participants.
Supporting optional participants is a relevant concept
for web applications, particularly for browser-side endpoints.

@@

\paragraph{Transport Abstraction}
The MPST-based web development framework presented by 
King et al. \cite{PureScript2019} defines the session runtime
in a way that parameterises the transportation mechanism.
The developer can run the session over different transportation,
provided that the custom transport implements the behaviour
of the \texttt{Transport} type class.

@@

%\paragraph{EFSM Encoding as Typestates}
%The concept of \textit{typestates} define the interface of
%an object to be dependent on its private state, meaning
%it can change at runtime.
%This is compatible with the EFSM abstraction from the
%MPST framework: the permitted methods (i.e. channel actions)
%of the EFSM depends on its current state.
%In fact, Kouzapas et al. \cite{StMungo} presents a way to
%generate typestate specifications for each endpoint of a
%Scribble protocol; Gay et al. \cite{ModularST} proposes
%an encoding of session types in object-oriented languages
%and discuss their approach with respect to typestates.
%
%Future work can explore implementing typestates
%in TypeScript. A linear type system would be useful towards 
%implementing typestates, but this is lacking in TypeScript.
%One interesting possibility for approximating this would
%be to combine \textit{decorators} with \textit{transformers}:
%the former lets the developer define typestate-related metadata,
%whilst the latter can parse the metadata and
%transform the source code into another TypeScript program that acts
%as an ``intermediate representation'',
%such that only implementations that respect the typestate specification
%will type-check with its intermediate representation.

### 2020-06-10 report/sections/evalconcl/eval.tex
Report: complete conclusion

\section{Binary Sessions: \fancyname{Calculator}}
\label{section:evalcalculator}
\begin{itemize}
\item sample client implementation + snippet of generated code (full in appendix)
\item show that not all the UI is being controlled by the runtime (i.e. calculator app has a header/fake navbar)
\item show the calculator logic achievable using React Context
\item sample server implementation + snippet of generated code (full in appendix)
\end{itemize}

\begin{figure}[!ht]
\begin{lstlisting}[language=Scribble, tabsize=2]
module BinarySession;

global protocol Calculator(role Svr, role Client) {
  choice at Client {
    Add(number, number) from Client to Svr;
    Res(number)         from Svr to Client;
    do Calculator(Svr, Client);
  } or {
    Mult(number, number) from Client to Svr;
    Res(number)          from Svr to Client;
    do Calculator(Svr, Client);
  } or {
    Quit()               from Client to Svr;
    Terminate()          from Svr to Client;
  }
}
\end{lstlisting}
\captionof{lstlisting}{The \tprotocol{Calculator} Protocol}
\label{lst:calculator}
\end{figure}

\section{Multiparty Sessions: \fancyname{Noughts and Crosses}}
\label{section:evalgame}
\begin{itemize}
\item sample client implementation -- show the flexibility of factory API for sending to be used in the UI part (since move selection can be be made )
\end{itemize}

>>>>

\section{Multiparty Sessions: \fancyname{Noughts and Crosses}}
\label{section:evalgame}
\begin{itemize}
\item sample client implementation -- show the flexibility of factory API for sending to be used in the UI part (since move selection can be be made )
\item show that not all the UI is being controlled by the runtime (i.e. calculator app has a header/fake navbar)
\item show the calculator logic achievable using React Context
\item sample server implementation + snippet of generated code (full in appendix)
\end{itemize}
